---
title: "Merging Datasets"
name: "Hazel Lutz"
date: "March 16, 2026"
output: 
  html_document:
    theme: sandstone
    toc: true
    toc_float: true
    number_sections: true
    code_folding: hide
---

```{r packages}
lapply(c("readr", "dplyr", "tidyr", "haven", "ggplot2"), library, character.only = TRUE)
# haven is for dealing with stata format
```

First, just importing from all of my CSVs.
```{r importing}
pruned_dir = "./data/pruned/"
intl_cow = read_csv(paste(pruned_dir, "intnl_filtered_cow.csv", sep=""))
nonintl_cow = read_csv(paste(pruned_dir, "nonintnl_filtered_cow.csv", sep=""))
vdem = read_csv(paste(pruned_dir, "vdem.csv", sep=""))
afro_old = read_csv(paste(pruned_dir, "afro.csv", sep=""))
americas_old = read_csv(paste(pruned_dir, "americas.csv", sep=""))
arab_old = read_csv(paste(pruned_dir, "arab.csv", sep=""))
```

```{r exploring_pub_op}
colnames(afro_old)
colnames(americas_old)

afro = afro_old %>%
	rename(
	       country = country,
	       urbrur = urbrur,
	       age = q1a,
	       country_dir = q3a,
	       economy_eval = q4a,
	       economic_improve = q5a,
	       anticipated_economy = q5b,
	       felt_safety = q7a,
	       felt_free_speech = q9a,
	       felt_free_org = q9b,
	       felt_free_voting = q9c,
	       felt_election_freedom = q14a,
	       fear_political_violence = q14c,
	       support_dem = q23,
	       felt_democracy = q31,
	       felt_dem_change = q32a,
anticipated_dem_change = q32b,
		democracy_over_efficacy = q33,
		role_of_opposition = q37,
	       most_important_issue1 = q45pt1,
	       most_important_issue2 = q45pt2,
most_important_issue3 = q45pt3,
	       gender_violence_taken_seriously = q53b,
	       neighbors_diff_party = q87e,
	       marry_diff_party = q88a,
	       mutual_citizenship = q88b
	) %>%
	mutate(economic_improve = if_else(economic_improve %in% c(1:5), as.character(economic_improve), NA_character_))
```

```{r}
americas = americas_old %>% 
	rename(
	       country = pais,
	       wave = wave,
	       year = year,
	       idnum = idnum,
	       province = prov,
	       urbrur = ur,
	       most_important_issue1 = a4,
	       economic_improve = idio2,
	       community_trust = it1,
	       respect_political_institutions = b2,
	       rights_respected = b3,
	       pride_political_system = b4,
	       support_political_system = b6,
	       trust_in_elections = b47a,
	       presidential_performance = m1,
	       trust_in_govt = anestg,
	       democracy_performance = pn4,
	       gender_violence_taken_seriously = aojg2n
	) %>%
	mutate(economic_improve = if_else(economic_improve %in% c(1:3), as.character(economic_improve), NA_character_))
# DATE COLUMN NON EXISTENT FOR WAVE 1? ARE YOU JOKING
```

```{r}
arab = arab_old %>%
	rename(
	       urbrur = q13,
	       most_important_issue1 = q2061a,
	       current_economy = q101,
	       anticipated_economy = q102,
	       community_trust = q103,
	       economic_improve = q114,
	       trust_in_govt = q201a_1,
	       trust_in_parliament = q201a_3,
	       trust_local_govt = q201a_5,
	       trust_pres = q201a_31b,
	       trust_pm = q201a_31c,
	       trust_army = q201b_6,
	       trust_police = q201b_4,
	       felt_safety = q105a,
	       felt_responsiveness = q276,
	       felt_free_expression = q521_1,
	       felt_improvement_prev_generation = q867a,
	       felt_improvement_next_generation = q867b,
	       felt_democracy = q511
	       ) %>%
		mutate(economic_improve = if_else(economic_improve %in% c(1:3), as.character(economic_improve), NA_character_))
```

```{r oil}
oil = read_csv("./data/pruned/oil.csv")
oil_prep = oil %>%
	arrange(country, year) %>%
	group_by(country) 
```

```{r merge_for_now}
intl_merged = intl_cow %>%
	pivot_longer(
		     cols = matches("^(Start|End)(Mo|Dy|Yr)\\d+$"),
		     names_to = c(".value", "episode"),
		     names_pattern = "([A-Za-z]+)(\\d+)"
	) %>%
	filter(EndYr != -8) %>%
	mutate(episode = as.numeric(episode)) %>%
	group_by(WarNum) %>%
	mutate(is_last_endyr = episode == max(episode)) %>%
	ungroup()
intl_merged_filtered = intl_merged %>%
	select(WarNum, WarName, SideA, episode, StartYr, EndYr, is_last_endyr, TotalBDeaths)

nonintl_merged = nonintl_cow %>%
	pivot_longer(
		     cols = matches("^(Start|End)(Mo|Dy|Yr)\\d+$"),
		     names_to = c(".value", "episode"),
		     names_pattern = "([A-Za-z]+)(\\d+)"
	) %>%
	filter(EndYr != -8) %>%
	mutate(episode = as.numeric(episode)) %>%
	group_by(WarNum) %>%
	mutate(is_last_endyr = episode == max(episode)) %>%
	ungroup()
nonintl_merged_filtered = nonintl_merged %>%
	select(WarNum, WarName, SideA, episode, StartYr, EndYr, is_last_endyr, TotalBDeaths)
nonintl_merged_filtered

vdem_prep = vdem %>%
	arrange(country_name, year) %>%
	group_by(country_name) %>%
	mutate(
	       gdppc_next = lead(e_gdppc, n=1, order_by = year),
	       partipdem_next = lead(v2x_partipdem, n=1, order_by = year),
	       gdppc_diff = gdppc_next - e_gdppc,
	       partipdem_diff = partipdem_next - v2x_partipdem
	) %>%
	ungroup() %>%
	select(country_name, year, gdppc_diff, partipdem_diff)

full_intl = intl_merged_filtered %>%
	left_join(
		  vdem_prep,
		  by = c("SideA" = "country_name", "EndYr" = "year")
	) %>%
	left_join(
		  oil_prep,
		  by = c("SideA" = "country", "EndYr" = "year")
	) %>%
	select(WarNum, WarName, SideA, episode, EndYr, is_last_endyr, TotalBDeaths, gdppc_diff, partipdem_diff, oil_rents)

full_nonintl = nonintl_merged_filtered %>%
	left_join(
		  vdem_prep,
		  by = c("SideA" = "country_name", "EndYr" = "year")
	) %>%
	left_join(
		  oil_prep,
		  by = c("SideA" = "country", "EndYr" = "year")
	) %>%
	select(WarNum, WarName, SideA, episode, EndYr, is_last_endyr, TotalBDeaths, gdppc_diff, partipdem_diff, oil_rents)

write_csv(full_intl, "./data/pruned/intl.csv")
write_csv(full_nonintl, "./data/pruned/nonintl.csv")
```

QCA cols:
* previous economic improvement (check scales) (fallback to other indicators)
  	* fuzzy set: full membership if moderate improvement, full nonmembership at slight deterioration, crossover point at "staying the same"
* community trust (no replacement)
  	* full membership at any form of agreeing, full nonmembership at any form of disagreeing, crossover point around neutral
* differential between trust in national and local government 
  	* generally, full membership when national is at least that of local, full nonmembership when natinal is substantially worse than local, crossover point when slightly worse
* democratic performance (partipdem) (v-dem)
  	* fuzzy set based on descriptive statistics most likely (maybe focus on increase from previous year)?
* petroleum/coal/natural gas 
  	* fuzzy set based on descriptive statistics of vdem e_total_fuel_income_pc
* total deaths in wartime
  	* fuzzy set -- can probably get away with descriptive statistics?
* peacekeeping operations (would be good but would have to be done manually -- perhaps olsen has that coding done already?)
  	* crisp set

QCA cols:
* previous economic improvement (check scales) (fallback to other indicators)
  	* fuzzy set: full membership if moderate improvement, full nonmembership at slight deterioration, crossover point at "staying the same"
* differential between trust in national and local government (any other ways to check this?)
  	* generally, full membership when national is at least that of local, full nonmembership when natinal is substantially worse than local, crossover point when slightly worse
* democratic performance (partipdem) (v-dem)
  	* fuzzy set based on descriptive statistics most likely (maybe focus on increase from previous year)?
* petroleum/coal/natural gas (v-dem)
  	* fuzzy set based on descriptive statistics of vdem e_total_fuel_income_pc
* total deaths in wartime (cow)
  	* fuzzy set -- can probably get away with descriptive statistics?
* peacekeeping operations (would be good but would have to be done manually -- perhaps olsen has that coding done already?)
  	* crisp set




