---
title: "R Does What Py Don't"
name: "Hazel Lutz"
date: "February 8, 2026"
output: 
  html_document:
    theme: sandstone
    toc: true
    toc_float: true
    number_sections: true
    code_folding: hide
---

First, we'll install some packages that we'll be using for this, as well as EDA later (I'll adapt this file when the time comes).
```{r packages}
lapply(c("readr", "dplyr", "haven", "ggplot2"), library, character.only = TRUE)
# haven is for dealing with stata format
```

```{r arabbarometer_cols}
arab_dir = "./data/arabbarometer/alldata/arab" # for loop
arab6_dir = paste(arab_dir, "6_", sep="")
colnames_list = vector("list", 3)
for (i in 1:3) {
	current_csv_dir = paste(arab6_dir, i, ".csv", sep="")
	colnames_list[[i]] = colnames(read_csv(current_csv_dir))
}
colnames_list
# cols in first part not in second
setdiff(colnames_list[[1]], colnames_list[[2]])
# cols in second part not in first
setdiff(coln
```

```{r arabbarometer6_combine}
data_dir = "./data/"
arab_dir = "./data/arabbarometer/alldata/arab" # for loop
arab6_dir = paste(arab_dir, "6_", sep="")
arab6_df = data.frame()
for (i in 1:3) {
	current_csv_dir = paste(arab6_dir, i, ".csv", sep="")
	current_df = read_csv(current_csv_dir)
	arab6_df = bind_rows(arab6_df, current_df)
}
write_csv(arab6_df, "./data/arabbarometer/alldata/arab6.csv")
```

```{r arabbarometer}
df_list = list()
arab = data.frame()
for (i in 1:8) {
	current_csv_dir = paste(arab_dir, i, ".csv", sep="")
	df_list[[i]] = read_csv(current_csv_dir) %>% 
		rename_with(tolower) %>% 
		mutate(across(everything(), as.character)) # ty stackexchange for the fix yayyyy
}

arab = bind_rows(df_list) 
write.csv(arab, "./data/pruned/arab.csv")
```


```{r afrobarometer}
df_list = list()
afro_dir = "./data/afrobarometer/alldata/afro"
afro = data.frame()
for (i in 1:9) {
	current_sav_dir = paste(afro_dir, i, ".sav", sep="")
	df_list[[i]] = read_sav(current_sav_dir, encoding = "latin1") %>%  # issue with special characters and whatnot
		rename_with(tolower) %>% # lowercases col names
		mutate(across(everything(), as.character))
}
afro = bind_rows(df_list)
write.csv(afro, "./data/pruned/afro.csv")
```

pais = country
wave = wave of survey
year = year of survey
idnum = unique id
prov/prov1t = province
ur = 1: city, 2: suburbs/outskirts of city, 3: town near rural, 4: rural
a4: most serious problem
idio2: economic situation compared to 12 mos ago is 1: better, 2: same, 3: worse
it1: community trust 1: very, 2: somewhat, 3: somewhat untrust,, 4: very untrust
(1-7, not at all - a lot)
b2: respect for political institutions of country
b3: citizens'b asic rights protected by political system
b4: pride of living under political system of country
b6: how much should one support political system of country
b47a: trust in elections
m1: performance of president
anestg: trust in national govt to "do what is right" (1-4, a lot to not at all)
pn4: 1-4, very satisfied, satisfied, dissatisfied, very dissatisfied with the way democracy works in country
aojg2n: likelihood of police to take seriously a woman being beaten by her partner (1: very likely, 2: somewhat likely, 3: a little likely, 4: not at all likely)
mil10a: trust in China (1: very trustworth, 4: not at all trustworthy)
mil10e: same for America

```{r americasbarometer}
americas_dir = "./data/americasbarometer/americas.dta"
americas_cols = c("pais", "wave", "year", "idnum", "prov", "prov1t", "ur", "a4", "idio2", "it1", "b2", "b3", "b4", "b6", "b47a", "m1", "anestg", "pn4", "aojg2n", "mil10a", "mil10e")
americas = read_dta(americas_dir, col_select = americas_cols)
write_csv(americas, "./data/pruned/americas.csv")
```
