---
title: "R Does What Py Don't"
name: "Hazel Lutz"
date: "February 8, 2026"
output: 
  html_document:
    theme: sandstone
    toc: true
    toc_float: true
    number_sections: true
    code_folding: hide
---

First, we'll install some packages that we'll be using for this, as well as EDA later (I'll adapt this file when the time comes).
```{r packages}
lapply(c("readr", "dplyr", "haven", "ggplot2"), library, character.only = TRUE)
# haven is for dealing with stata format
```

For ArabBarometer, there were three different segments to one survey wave, and so this was to check for discrepancies between them before opting to just include all columns to make one CSV, and then deal with any issues from there.
```{r arabbarometer_cols}
arab_dir = "./data/arabbarometer/alldata/arab" # for loop
arab6_dir = paste(arab_dir, "6_", sep="")
colnames_list = vector("list", 3)
for (i in 1:3) {
	current_csv_dir = paste(arab6_dir, i, ".csv", sep="")
	colnames_list[[i]] = colnames(read_csv(current_csv_dir))
}
colnames_list
# cols in first part not in second
setdiff(colnames_list[[1]], colnames_list[[2]])
```

Putting together the ArabBarometer Wave 6 CSVs
```{r arabbarometer6_combine}
data_dir = "./data/"
arab_dir = "./data/arabbarometer/alldata/arab" # for loop
arab6_dir = paste(arab_dir, "6_", sep="")
arab6_df = data.frame()
for (i in 1:3) {
	current_csv_dir = paste(arab6_dir, i, ".csv", sep="")
	current_df = read_csv(current_csv_dir)
	arab6_df = bind_rows(arab6_df, current_df)
}
write_csv(arab6_df, "./data/arabbarometer/alldata/arab6.csv")
```
ID
DATE
Q13: 1 = urban, 2 = rural, 3= camp
Q2061A: most important issue
Q101: current economic situation in country: 1= very good, 2=good, 3=bad, 4=very bad
Q102: anticipated economic in next few years in country: 1 much better, 2 somewhat better, 3 same, 4 somewhat worse, 5 much worse
Q103: trust others (1) or be careful with others (2)
trust in institutions (1 a great deal, 2 quite a lot, 3 not a lot, 4 none at all)
Q201A_1: government
Q201A_3 parliament
Q201A_5: local government
Q201A_31B: president/head of state
Q201A_31C: prime minister/had of government
Q201B_6: army
Q201B_4: police
Q105A: safety of neighborhood (1 very safe, 2 somewhat, 3 somewhat, 4 very unsafe)
Q276: felt responsiveness of government (1 very, 2 largely, 3 not vary, 4 not at all)
Q521_1: felt freedom to express opinions (1 guaranteed greatly, 2 guaranteed mediumly, 3 guaranteed limitedly, 4 not guaranteed at all)
Q511: felt democracy (0-10, none to as democratic as possible)
Q867A: my life in relation to parents' lives (1 better, 2 same, 3 worse)
Q867B: same but for if children's will (1 better, 2 same, 3 worse)




And then compiling all of those into one.
```{r arabbarometer}
df_list = list()
arab_cols = c("id", "wave", "dateintr", "country", "q13", "q2061a", "q101", "q102", "q103", "q114", "q201a_1", "q201a_3", "q201a_5", "q201a_31b", "q201a_31c", "q201b_6", "q201b_4", "q105a", "q276", "q521_1", "q511", "q867a", "q867b")

wave_dates = c("2006-2009", "2010-2011", "2012-2014", "2016-2017", "2018-2019", "July 2020 - April 2021")
arab = data.frame()
for (i in 1:8) {
	current_csv_dir = paste(arab_dir, i, ".csv", sep="")
	current_df = read_csv(current_csv_dir) %>%
		rename_with(tolower)

	# this next chunk is to deal with renaming the date column since R interprets it as a function in case there is no date column to begin with which is dealt with later. sigh 
	if (!"dateintr" %in% colnames(current_df)) {
		if ("date" %in% colnames(current_df)) {
			current_df = current_df %>% rename(dateintr = date)
		} else {
			current_df = current_df %>% mutate(dateintr = NA_character_)
		}
	}
	df_list[[i]] = current_df %>% 
		rename_with(tolower) %>% 
		mutate(wave = as.character(i)) %>%
		mutate(dateintr = if (i <= 6) wave_dates[i] else as.character(dateintr)) %>%
		mutate(across(everything(), as.character)) # ty stackexchange for the fix yayyyy
}
arab = bind_rows(df_list) %>%
	select(arab_cols)
colnames(arab)
write.csv(arab, "./data/pruned/arab.csv")
```

AfroBarometer was fairly similar with fewer issues.
country: country
urbrur: urban vs. rural (1 urban, 2 rural, 3 peri-urban)
dateintr: date of interview 
q1: age
q3: direction of country (1 wrong, 2 right)
q4a: economic condition (1 very bad, 5 very good)
q5a: economic improvement in past 12 months (1 much worse, 5 much better)
q5b: anticipated economic conditions (1 much worse, 5 much beter)
q7a: felt unsafe in neighborhood (0 never, 4 always)
q9a: felt free speech (1 none, 4 completely)
q9b: felt free organization (1 none, 4 completely)
q9c: felt free voting (1 none, 4 completely)
q14a: felt freeness of previous election (1 none, 4 completely) 
q14c: fear of political violence (0 a lot, 3 not at all)
q23: support for democracy (1 doesn't matter, 2, sometimes preferable, 3 always preferable)
q30: felt democracy (1 not a democracy, 4 full democracy)
q31: satisfaction with democracy (0 not a democracy, 1 not satisfied, 4 very satisfied)
q32a: democratic change compared to five years ago (1 much less demo, 5 much more demo)
q32b: anticipated democratic change in five years (1 much less, 5 much more) 
q33a: does competition between political parties lead to violent conflict? (0 never, 3 always)
q33e: unequal treatment under law (0 never, 3 always)q37a: trust president (0 not at all, 3 a lot)
q37b: trust parliament ("")
q37d: trust local government ("")
q37e: trust ruling party ("")
q37f: trust opposition parties ("")
q37g: trust police ("")
q37h: trust army ("")
q45pt1: most important issue *
q45pt2 (and pt3): other choices to above
q53b: likelihood of women reporting gender-based violence being taken seriously by police (1 very unlikely, 4 very likely)
q87e: fine having neighbors of diff political party (1 strongly dislike, 5 strongly like, 3 would not care) *
q88a: family member marrying outside ethnic group (1 strongly dislike, 5 strongly like, 3 would not care) *
a88b: others see respondent as citizen like them (q88b)



```{r afrobarometer}
df_list = list()
afro_dir = "./data/afrobarometer/alldata/afro"
afro = data.frame()
afro_cols = c("country", "urbrur", "wave", "dateintr", "q1a", "q1b", "q3a", "q3b", "q4a", "q5a", "q5b", "q7a", "q9a", "q9b", "q9c", "q14", "q14a", "q14c", "q23", "q30", "q31", "q32", "q33", "q37", "q44pt1", "q44pt2", "q44pt3", "q45pt1", "q45pt2", "q45pt3", "q53b", "q88", "q32a", "q32b", "q88a", "q88b", "q87e")
for (i in 1:9) {
	current_sav_dir = paste(afro_dir, i, ".sav", sep="")
	df_list[[as.character(i)]] = read_sav(current_sav_dir, encoding = "latin1") %>%  # issue with special characters and whatnot
		rename_with(tolower) %>% # lowercases col names (in case of some caps, some lower)
		mutate(across(everything(), as.character))
}
# puts together these dfs and makes a wave col based on name of index in df_list
afro = bind_rows(df_list, .id = "wave") %>%
		# sets date based on wave
		mutate(dateintr = if_else(wave == "1" & is.na(dateintr), "1999-2001", dateintr)) %>%
	select(afro_cols)

colnames(afro)
write.csv(afro, "./data/pruned/afro.csv")
```


pais = country
wave = wave of survey
year = year of survey
idnum = unique id
prov/prov1t = province
ur = 1: city, 2: suburbs/outskirts of city, 3: town near rural, 4: rural
a4: most serious problem
idio2: economic situation compared to 12 mos ago is 1: better, 2: same, 3: worse
it1: community trust 1: very, 2: somewhat, 3: somewhat untrust,, 4: very untrust
(1-7, not at all - a lot)
b2: respect for political institutions of country
b3: citizens'b asic rights protected by political system
b4: pride of living under political system of country
b6: how much should one support political system of country
b47a: trust in elections
m1: performance of president
anestg: trust in national govt to "do what is right" (1-4, a lot to not at all)
pn4: 1-4, very satisfied, satisfied, dissatisfied, very dissatisfied with the way democracy works in country
aojg2n: likelihood of police to take seriously a woman being beaten by her partner (1: very likely, 2: somewhat likely, 3: a little likely, 4: not at all likely)
mil10a: trust in China (1: very trustworth, 4: not at all trustworthy)
mil10e: same for America

```{r americasbarometer}
americas_dir = "./data/americasbarometer/americas.dta"
americas_cols = c("pais", "wave", "year", "idnum", "prov", "prov1t", "ur", "a4", "idio2", "it1", "b2", "b3", "b4", "b6", "b47a", "m1", "anestg", "pn4", "aojg2n", "mil10a", "mil10e")
americas = read_dta(americas_dir, col_select = americas_cols)
write_csv(americas, "./data/pruned/americas.csv")
```

Now for the conflict data. The conflict type dataset did not provide me with everything I was looking for, which means I may have to resort to a different database for this. On the search for one as I type. More specifically, it just doesn't have enough data for what I'm looking for, andit also doesn't classify by external intervention in the way I would (will elaborate more in properly typing it up).
```{r ucdp_conflicts}
ucdp_conflict_df = read_csv("./data/UcdpPrioConflict_v25_1.csv")  # this is for the list of valid conflicts
colnames(ucdp_conflict_df)
head(ucdp_conflict_df)
unique(ucdp_conflict_df$type_of_conflict)  # focused on type_of_conflict == 3 (intrastate, no external inv) and type_of_conflict == 4 (intrastate, external inv)
ucdp_conflict_df$ep_end_date = as.POSIXct(ucdp_conflict_df$ep_end_date)
cutoff_date = as.POSIXct("2002-01-01")
conflicts_no_ext_df = ucdp_conflict_df %>%
	filter(ep_end_date >= cutoff_date, type_of_conflict == 3, intensity_level == 2)  # intensity_level == 2 is war
conflicts_ext_df = ucdp_conflict_df %>%
	filter(ep_end_date >= cutoff_date, type_of_conflict == 4, intensity_level == 2)
write_csv(conflicts_no_ext_df, "./data/pruned/conflicts_no_ext.csv")  # these are both giving rather small datasets unfortunately
write_csv(conflicts_ext_df, "./data/pruned/conflicts_ext.csv")
```

Instead, I'm pivoting to the war database being used to Correlates of War, which has less recent data but has a more comprehensive list of conflicts. Will have to compare these two datasets to extract the best results.
```{r ucdp_ged}
cow_df = read_csv("./data/Intra-State-Wars-v5.1/INTRA-STATE WARS v5.1 CSV.csv")  
colnames(cow_df)
head(cow_df)
nonintnl_filtered_cow_df = cow_df %>%
	filter(EndYr1 >= 1990, Intnl == 0)
write_csv(nonintnl_filtered_cow_df, "./data/pruned/nonintnl_filtered_cow.csv")
```


```{r selecting_survey_cols}
afro_df = read_csv("./data/pruned/afro.csv")
arab_df = read_csv("./data/pruned/arab.csv")
colnames(arab_df)
colnames(afro_df)
```

```{r vdem}
vdem = read_csv("./data/V-Dem-CY-FullOthers-v16_csv/V-Dem-CY-Full+Others-v16.csv")
vdem_filtered = vdem %>%
	select(country_name, year, v2x_polyarchy, v2x_partipdem, v2x_egaldem, v2x_accountability, e_peaveduc, e_gdppc, e_total_fuel_income_pc)
# these are: electoral dem, participatory dem, egalitarian dem, accountability of government, education, gdp per captia, petroleum/coal/natural gas wealth)
write_csv(vdem_filtered, "./data/pruned/vdem.csv")
```
